---
title: "Proyecto2"
author: "Nosequiensoy007"
date: "12/11/2020"
output: html_document
---
# set libraries

```{r}
library(dplyr)
library(ggplot2)
library(summarytools)
library(cowplot)
```


#After downloadzip, file, unzip and then change wd

```{r}
setwd("C:/Users/juanfidel18/Desktop/Coursera/ReproducibleResearch")
```

# reading and Procesing Data

```{r}
stormdata <- read.csv("repdata_data_StormData.csv")
```

# Check dimensions and structure
```{r}
str(stormdata)
```

```{r}
freq(stormdata$EVTYPE, order = "freq", rows = 1:10) 
```
```{r}
summary(stormdata$INJURIES)
```
```{r}
summary(stormdata$FATALITIES)
```

```{r}
summary(stormdata$PROPDMG)
```

```{r}
freq(stormdata$PROPDMGEXP)
```
```{r}
summary(stormdata$CROPDMG)
```

```{r}
freq(stormdata$CROPDMGEXP)
```

##Data Processing

```{r}
stormpro <- select(stormdata, EVTYPE, FATALITIES, INJURIES, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP)
```

#now we can transform
```{r}
stormpro$EVTYPE <- as.character(stormpro$EVTYPE) 
stormpro$EVTYPE <- toupper(stormpro$EVTYPE)
```
```{r}
stormpro$PROPDMGEXP <- toupper(stormpro$PROPDMGEXP)
stormpro <- mutate(stormpro, propdmgexp2 = recode(PROPDMGEXP,
                                             "1" = 10,
                                             "2" = 100,
                                             "3" = 1000,
                                             "4" = 10000,
                                             "5" = 100000,
                                             "6" = 1000000,
                                             "7" = 10000000,
                                             "8" = 100000000,
                                             "H" = 100,
                                             "K" = 1000,
                                             "M" = 1000000,
                                             "B" = 1000000000,
                                             .default = 1)) 
stormpro <- mutate(stormpro, propdmgvalue = PROPDMG*propdmgexp2)
```
#calculate damage 

```{r}
stormpro$CROPDMGEXP <- toupper(stormpro$CROPDMGEXP)
stormpro <- mutate(stormpro, cropdmgexp2 = recode(CROPDMGEXP,  
                                                  "1" = 10,
                                                  "2" = 100,
                                                  "3" = 1000,
                                                  "4" = 10000,
                                                  "5" = 100000,
                                                  "6" = 1000000,
                                                  "7" = 10000000,
                                                  "8" = 100000000,
                                                  "H" = 100,
                                                  "K" = 1000,
                                                  "M" = 1000000,
                                                  "B" = 1000000000,
                                                 .default = 1))
stormpro <- mutate(stormpro, cropdmgvalue = CROPDMG*cropdmgexp2)
```
#subset injuries info

```{r}
injuries <- aggregate(INJURIES~EVTYPE, data = stormpro, FUN = sum) 
injuries <- injuries[order(injuries$INJURIES, decreasing = TRUE), ] 
injuries <- injuries[1:5, ] 
```
#fatality data by event


```{r}
fatalities <- aggregate(FATALITIES~EVTYPE, data = stormpro, FUN = sum) 
fatalities <- fatalities[order(fatalities$FATALITIES, decreasing = TRUE), ] 
fatalities <- fatalities[1:5, ]
```
#Properly damage by type
```{r}
propdamage <- aggregate(propdmgvalue~EVTYPE, data = stormpro, FUN = sum)
propdamage <- propdamage[order(propdamage$propdmgvalue, decreasing = TRUE), ]
propdamage <- propdamage[1:5, ]
```

```{r}
cropdamage <- aggregate(cropdmgvalue~EVTYPE, data = stormpro, FUN = sum)
cropdamage <- cropdamage[order(cropdamage$cropdmgvalue, decreasing = TRUE), ]
cropdamage <- cropdamage[1:5, ]
```

#Results
1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

```{r}
injgraph <- ggplot(data = injuries, aes(x = EVTYPE, y = INJURIES)) + labs(x="Event", y="Injuries") + geom_col() 
fatgraph <- ggplot(data = fatalities, aes(x = EVTYPE, y = FATALITIES)) + labs(x="Event", y="Fatalities") + geom_col()
title <- ggdraw() + draw_label("Most harmful events to population health", fontface = 'bold') + theme(plot.margin = margin(0,0,0,7))
plot_col <- plot_grid(injgraph,fatgraph, ncol = 1)
```
#plot
```{r}
plot_grid(title,plot_col, ncol = 1, rel_heights = c(0.1,1))
```
#results of question 1
#Tornados are the most harmful to population health, causing the most injuries and fatalities(33.2%).


#now go for question 2/ 
2. Across the United States, which types of events have the greatest economic consequences?

```{r}
propdmg  <- ggplot(data = propdamage, aes(x = EVTYPE, y = propdmgvalue)) + labs(x="Event", y="Property Damage") + geom_col() 
cropdmg  <- ggplot(data = cropdamage, aes(x = EVTYPE, y = cropdmgvalue)) + labs(x="Event", y="Crop Damage") + geom_col()
title2 <- ggdraw() + draw_label("Most harmful events to the economy", fontface = 'bold') + theme(plot.margin = margin(0,0,0,7))
plot_col2 <- plot_grid(propdmg,cropdmg, ncol = 1)
```

```{r}
plot_grid(title2,plot_col2, ncol = 1, rel_heights = c(0.1,1))
```

#results of question 2

#Floods cause the most property damage while droughts cause the most crop damage (37.1%)


